# IODS Final assignment
## Tina Lauronen 6.3.2017
### tina.lauronen@helsinki.fi

## Abstract [write this last!]

## [Brief description of the study]

This is the final assignment of the course *Introduction to Open Data Science*, University of Helsinki, spring 2017. In this assaignment I use the data of the research project *Cultural Distinctions, Generations and Change*. I am doing my doctoral studies in this project, shortly called [CuDiGE](http://blogs.helsinki.fi/cudige-project/).

In this assignment...

## The data

The data is collected from the cultural sections of European newspapers from 1960 to 2010. The newspapers are *ABC* and *El País* (hereafter *ABC/EP*) from Spain, *Dagens Nyheter* (*DN*) from Sweden, *The Guardian* (*GU*) from UK, *Helsingin Sanomat* (*HS*) from Finland, *Le Monde* (*LM*) from France and *Milliyet* (*MIL*) from Turkey. There are three randomly sampled weeks collected from each newspaper from ten-year intervals: 1960, 1970, 1980, 1990, 2000 and 2010. 

The data is coded by eleven coders using ATLAS.ti software. The unit of analysis is individual article and the total N = 13,161. The coding system included more than 300 codes that covered 41 variables.

The data has been saved from SPSS to MS Office Excel and then modified in RSudio. Originally the data consisted of 13,161 observations and 373 variables (including several recodings). After selecting variables for these analysis and removing the observations with missing variables the data set includes 6,506 observations and nine variables. The code is available in this [RScript](https://github.com/tinalauronen/IODS-final/blob/master/Data/creating_cudige.R).


## [Description of the data and the variables]

The CuDiGe data is collected and coded in addition to investigate cultural hierarchies and valuations and their change in time. The coding system covered some technical issues like the size of the article, but mainly concentrated to the contents like the cultural area or art form discussed in the article. In these analysis we use variables related to the art form, 
- what does the data relate to
- what do the variables present
- what has been done to the data

```{r, echo=FALSE, message=FALSE}
library(GGally)
```

```{r, echo = FALSE}
cudige <- read.csv("/Users/tlaurone/GitHub/tinalauronen/IODS-final/Data/cudige.csv")
str(cudige)
```

## [Distributions and relationships of the variables]
- visualizations

```{r, echo=FALSE}
counts <- table(cudige$Newspaper, cudige$Year)
barplot(counts, main = "Articles by year and newspaper", legend = rownames(counts), beside = TRUE)
```

- interpretations

## [Method]

## [Results]
- summaries, visualizations
- interpretations
- validation analysis of the method

## [Conclusions and discussion]
