# IODS Final assignment

## Tina Lauronen 6.3.2017

### tina.lauronen@helsinki.fi

## Abstract [write this last!]

## [Brief description of the study]

This is the final assignment of the course *Introduction to Open Data Science*, University of Helsinki, spring 2017. In this assaignment I use the data of the research project *Cultural Distinctions, Generations and Change*. I am doing my doctoral studies in this project, shortly called CuDiGE.

In this assignment I test a new way to approach the familiar data: I pick several interesting variables and experiment a new method for me, multiple correspondece analysis (MCA). Previously in this project we have been using logistic regression and multinomial logistic regression. The themes we have been working with this far have been related to highbrow and popular culture, established and emerging cultural areas and cultural globalization (the last one being the focus of my dissertation). The publication list is available at the [project web site](http://blogs.helsinki.fi/cudige-project/). This time I mix the themes a little bit and let an unsupervised method test if it finds the same kind of issues we have find. Also I include some less used and somewhat questionable variables (dimensional variables, see the presentation of the variables above) in the analysis just to see how they act with more familiar variables.

The research questions of this assignment are quite wide: First I will try to find dimensions to support our earlier findings related to old and new; highbrow and popular; and domestic and foreign cultural products. Then I wish to find out how our aesthetic, commercial and political dimensions relate to those aspects. As background variables I keep the year and the newspaper in the analysis.

## The data

The data is collected from the cultural sections of European newspapers from 1960 to 2010. The newspapers are *ABC* and *El País* (hereafter *ABC/EP*) from Spain, *Dagens Nyheter* (*DN*) from Sweden, *The Guardian* (*GU*) from UK, *Helsingin Sanomat* (*HS*) from Finland, *Le Monde* (*LM*) from France and *Milliyet* (*MIL*) from Turkey. There are three randomly sampled weeks collected from each newspaper from ten-year intervals: 1960, 1970, 1980, 1990, 2000 and 2010. 

The data is coded by eleven coders using ATLAS.ti software. The unit of analysis is individual article and the total N = 13,161. The coding system included more than 300 codes that covered 41 variables.

The data has been saved from SPSS to MS Office Excel and then modified in RSudio. Originally the data consisted of 13,161 observations and 373 variables (including several recodings). After selecting variables for these analysis and removing the observations with missing variables the data set includes 6,506 observations and nine variables. The code is available in this [RScript](https://github.com/tinalauronen/IODS-final/blob/master/Data/creating_cudige.R).

## The variables

The CuDiGe data is collected and coded in addition to investigate cultural hierarchies and valuations and their change in time. The coding system covered some technical issues like the size of the article, but mainly concentrated to the contents like the cultural area or art form discussed in the article. In these analysis the variables (and their values) are the following:

- *Year*
    - 1960; 1970; 1980; 1990; 2000 and 2010
- *Newspaper*
    - ABC/EP; DN; GU; HS; LM and MIL
- *Art_form* referring to the primary cultural area discussed in the article
    - Literature; Classical music; Popular music; Film; TV; Theatre; Fine arts and Other
- *Artist_time* referring to the time horizon of the artist of piece of art discussed in the article
    - Present; Post WW2; Pre WW2 and Historical
- *Format* referring to the format of the piece of art
    - Live; Recording
- *Aesthetic* referring to the aesthetic dimension: explicit judgements on the good or poor quality of the piece of art treated in the article
     - Yes/No
- *Commercial* referring to the commercial dimension: explicit references to money or other economic indicators
     - Yes/No
- *Political* referring to the political dimension: explicit references to politics
     - Yes/No

```{r, echo=FALSE, message=FALSE}
library(GGally)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
library(corrplot)
library(FactoMineR)
```

```{r, echo = FALSE, message=FALSE}
cudige <- read.csv("/Users/tlaurone/GitHub/tinalauronen/IODS-final/Data/cudige.csv")
cudige$Year <- factor(cudige$Year, levels = c(1960, 1970, 1980, 1990, 2000, 2010), labels = c("1960", "1970", "1980", "1990", "2000", "2010"))
```

## Descriptive statistics

First we take a look at the distribution on the variables. In Figure 1 there are the frequencies of the variables presented in alphabetical order. 

```{r, echo=FALSE, warning=FALSE, fig.cap="**Figure 1:** Distributions of the variables"}
gather(cudige) %>% ggplot(aes(value)) + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 6)) + facet_wrap("key", scales = "free")
```

The amount of articles increases in time, although 2010 there is a slight decrease visible. There are clear differences among the newspapers as well, the smalles *MIL* includes here only a little more than 500 articles while the biggest *HS* includes almost 1,500 articles.

The most common art form is *literature*, followed by *popular* and *classical* music and *film*. The origin of the artist/piece of art is most commonly *domestic* and the time horizon of the artist/piece of art *present*. The pieces of art are more commonly recording than live performances.

Of the dimensional variables the most common is *aesthetic dimension*. There is aesthetic dimension coded in almost 4,000 articles, while *political* is found in less than 2,000 and *commercial* only in 1,000 articles.

The same results are available in the summary table.

```{r, echo=FALSE}
summary(cudige, maxsum = 8)
```


## Multiple Correspondence Analysis (MCA)

## Results

```{r, echo=FALSE, fig.height=8, fig.width=8}
mca <- MCA(cudige, graph = FALSE)
plot(mca, invisible = c("ind"), habillage = "quali")
```

- summaries, visualizations
- interpretations
- validation analysis of the method

## [Conclusions and discussion]
